Precision calculations for cosmic shear

### 2017-02-08 cfhtlens.tex
MK edits

In this work, we focus on the weak-lensing power spectrum projection, and
assess the impact of various approximations on cosmological constraints from
CFHTLenS. We re-analyse the 2D CFHTLenS data from \cite{CFHTLenS-2pt-notomo},
adopting the same priors and non-linear power spectrum corrections\footnote{As
an example of significant effects that need to be accounted for using the
revised non-linear power spectrum of \cite{2012ApJ...761..152T} in place of
\cite{2003MNRAS.341.1311S} results in a decrease of $0.6 \sigma$ with $\sigma_8
(\Omega_{\rm m}/0.27)^{0.6} =0.768^{+0.029}_{-0.031}$.} from
\cite{2003MNRAS.341.1311S} as the original \cite{CFHTLenS-2pt-notomo} analysis.
For a first-order standard Limber flat-sky approximation (L1Fl) we find
$\sigma_8 (\Omega_{\rm m}/0.27)^{0.6} =0.787^{+0.031}_{-0.033}$. For a
second-order extended Limber flat-sky approximation we find $\sigma_8
(\Omega_{\rm m}/0.27)^{0.6} =X^{+X}_{-X}$, a negligible change of
$X\sigma$.

@@

\ch{This text taken from the comment on paper; We note that
\citet{2016arXiv161104954K} use their `Flat' approximation for their `Full'
non-Limber analysis of the CFHTLenS data.   This is because on a sphere,
correlation functions cannot be related to the power spectrum by a Hankel
transform (e.g. equation~\ref{eqn:xiGG}) as the power spectrum is not defined
for non-integer $\ell$. In Figure~\ref{fig:Cl_xi}, however, we show that this
chosen version of a flat-sky approximation serves to exacerbate the tension in
theoretical models when a flat-sky solution is compared with the spherical-sky
solution in the Limber approximated case.  This is not the case when
considering the standard $T_\ell = 1$ flat-sky approximation that has been used
in the majority of cosmic shear studies to date.  We would therefore argue that
it is more appropriate to use a $T_\ell = 1$ flat-sky approximation when
carrying out a `Full' non-Limber analysis of the shear correlation function
obtained from the Hankel transformation (equation~\ref{eqn:xiGG}). We need to
take a decision here about what to use.  Do we need a Full Flat case in Table
1, and to demonstrate that Ext2Fl is very close to FullFl?  And then use Ext2Fl
as our "truth" for the correlation function?}

>>>>

In this work, we focus on the weak-lensing power spectrum projection, and
assess the impact of various approximations on cosmological constraints from
CFHTLenS. We re-analyse the 2D CFHTLenS data from \cite{CFHTLenS-2pt-notomo},
adopting the same priors and non-linear power spectrum corrections\footnote{As
an example of significant effects that need to be accounted for using the
revised non-linear power spectrum of \cite{2012ApJ...761..152T} in place of
\cite{2003MNRAS.341.1311S} results in a decrease of $0.6 \sigma$ with $\sigma_8
(\Omega_{\rm m}/0.27)^{0.6} =0.768^{+0.029}_{-0.031}$.} from
\cite{2003MNRAS.341.1311S} as the original \cite{CFHTLenS-2pt-notomo} analysis.
We use the real-space two-point correlation function of shear, which in flat-sky
coordinates is a Hankel transform of the shear power spectrum,
%
\begin{eqnarray}
  \xi_+(\theta) 
  % &
  = \frac 1 {2\pi} \int {\rm d} \ell \, \ell {\rm J}_0(\ell
   \theta)
  P^\gamma(\ell);
  \quad
  %\nonumber \\
   %
   \xi_-(\theta)
  %&
  = \frac 1 {2\pi} \int
   {\rm d} \ell \, \ell {\rm J}_4(\ell \theta)
  P^\gamma(\ell).
  %
   \label{eqn:xiGG}
\end{eqnarray}
%
As in \cite{CFHTLenS-2pt-notomo} we fit both components $\xi_+$ and $\xi_-$
between angular scales $\theta = 0.8$ and $350$ arc minutes, and use a $N$-body
simulation estimate of the non-Gaussian covariance including the
cross-covariance between both components. The Bayesian Monte-Carlo parameter
sampling is performed using the publically available software
\textsc{CosmoPMC}\footnote{\texttt{www.cosmopmc.info}} \citep{WK09,KWR10}. The
cosmological modelling part including the various lensing projections are
performed with the library
\textsc{nicaea}\footnote{\texttt{http://www.cosmostat.org/software/nicaea}}
\citep{KB09}. \mk{Check reference.}

For a first-order standard Limber flat-sky approximation (L1Fl) we find
$\sigma_8 (\Omega_{\rm m}/0.27)^{0.6} =0.787^{+0.031}_{-0.033}$, the same
result that was published in \cite{CFHTLenS-2pt-notomo}. Using the second-order
extended Limber flat-sky approximation results in $\sigma_8 (\Omega_{\rm
m}/0.27)^{0.6} =X^{+X}_{-X}$, a negligible change of $X\sigma$.

@@

\ch{This text taken from the comment on paper; We note that
\citet{2016arXiv161104954K} use their `Flat' approximation for their `Full'
non-Limber analysis of the CFHTLenS data. This is because on a sphere,
correlation functions cannot be related to the power spectrum by a Hankel
transform (\ref{eqn:xiGG}) as formally the power spectrum is not defined for
non-integer $\ell$. In Fig.~\ref{fig:Cl_cases}, however, we show that this
chosen version of a flat-sky approximation serves to exacerbate the tension in
theoretical models when a flat-sky solution is compared with the spherical-sky
solution in the Limber approximated case.  This is not the case when
considering the standard and extended hybrid flat-sky Limber approximation
(both for which $p(\ell) = 1$, which corresponds to $T_\ell$ in
\citet{2016arXiv161104954K}, see Sect.~\ref{sec:Limber_flat_sky}) that has been
used in the majority of cosmic shear studies to date.  We would therefore argue
that it is more appropriate to use a $p(\ell) = 1$ flat-sky approximation when
carrying out a `full' non-Limber analysis of the shear correlation function
obtained from the Hankel transformation (equation~\ref{eqn:xiGG}). We need to
take a decision here about what to use. Do we need a Full Flat case in Table 1,
and to demonstrate that Ext2Fl is very close to FullFl?  And then use Ext2Fl as
our "truth" for the correlation function?}

### 2017-02-08 discussion.tex
MK edits

\citet{CFHTLenS-2pt-notomo} present a detailed comparison of cosmological
constraints obtained from a range of different two-point shear statistics
including the shear correlation function, $\xi_\pm$, the aperture-mass
dispersion, $\langle M_ {\rm ap}^2 \rangle2$ \citep{1998MNRAS.296..873S}, and
the COSEBIs, $E_n$ \citep{COSEBIs}.  These statistics are linearly related to
the shear power spectrum in the flat-sky approximation via integrals of the form (omitting tomographic indices),
%
\begin{align}
  \langle M_ {\rm ap}^2 \rangle(\theta) & = \int_0^{\infty}\frac{\d \ell \ell}{2\pi}
  \hat U^2(\theta\ell) P^\gamma(\ell),
  \label{eqn:integ}
  \\
  E_n & = \int_0^{\infty}\frac{\d \ell \ell}{2\pi} W_n(\ell) P^\gamma(\ell),
\end{align}
%
where the Fourier-space filter functions $\hat U$ and $W_n$ are defined in
\cite{1998MNRAS.296..873S} and \cite{COSEBIs}. The corresponding equations for
$\xi_\pm$ are given in equation~\ref{eqn:xiGG}. Figure~\ref{fig:filters} shows
the integrands of these statistics for two cases, normalised to their maximum
value, where the integrands are of the form $\ell F(\ell) P^\gamma(\ell)$. For
$\xi_\pm$, $P^\gamma(\ell)$ is equal to the sum of the E and B-mode power spectra
which cannot easily be separated,
motivating the development of the aperture-mass dispersion statistic, $\langle
M_ {\rm ap}^2 \rangle$. This statistic is, however, a lossy conversion and is
biased by small angular separations, where blending of galaxies makes shear
measurement challenging \citep{KSE06}. The COSEBIs statistic tackles both
these shortcomings.  The upper two panels of figure~\ref{fig:filters} show the
integrands\footnote{Note that the lower left panel of Fig.~1 in
\citet{2016arXiv161104954K} shows an integrated form of this function for a
maximum angular separation of $100^\prime$. However, in
\citet{CFHTLenS-2pt-notomo}, the data used in \citet{2016arXiv161104954K},
$\theta$ is between $0.8'$ and $350'$.} of $\xi_\pm$ for $\theta=100'$ and
$\theta=350'$. The lower middle panel in Figure~\ref{fig:filters} shows the
COSEBIs integrands for two angular ranges, $[1',100']$ and $[0.8',350']$, where
we only show the integrands for the lowest COSEBIs mode, $E_1$, as the higher
modes generally probe larger $\ell$-modes.  Finally, the lowest panel shows the
integrands of aperture mass dispersion statistics, for the same two maximum
angular ranges.

>>>>

\citet{CFHTLenS-2pt-notomo} present a detailed comparison of cosmological
constraints obtained from a range of different two-point shear statistics
including the shear correlation function, $\xi_\pm$, the aperture-mass
dispersion, $\langle M_ {\rm ap}^2 \rangle$ \citep{1998MNRAS.296..873S}, and
the COSEBIs, $E_n$ \citep{COSEBIs}.  These statistics are linearly related to
the shear power spectrum in the flat-sky approximation via integrals of the
form (omitting tomographic indices),
%
\begin{align}
  \langle M_ {\rm ap}^2 \rangle(\theta) & = \int_0^{\infty}\frac{\d \ell \ell}{2\pi}
  \hat U^2(\theta\ell) P^\gamma(\ell),
  \label{eqn:integ}
  \\
  E_n & = \int_0^{\infty}\frac{\d \ell \ell}{2\pi} W_n(\ell) P^\gamma(\ell),
\end{align}
%
where the Fourier-space filter functions $\hat U$ and $W_n$ are defined in
\cite{1998MNRAS.296..873S} and \cite{COSEBIs}. The corresponding equations for
$\xi_\pm$ are given in equation~\ref{eqn:xiGG}. Figure~\ref{fig:filters} shows
the integrands of these statistics for two cases, normalised to their maximum
value, where the integrands are of the form $\ell F(\ell) P^\gamma(\ell)$. For
$\xi_\pm$, $P^\gamma(\ell)$ is equal to the sum of the E and B-mode power spectra
which cannot easily be separated,
motivating the development of the aperture-mass dispersion statistic, $\langle
M_ {\rm ap}^2 \rangle$. This statistic is, however, a lossy conversion and is
biased by small angular separations, where blending of galaxies makes shear
measurement challenging \citep{KSE06}. The COSEBIs statistic tackles both
these shortcomings.  The upper two panels of figure~\ref{fig:filters} show the
integrands\footnote{Note that the lower left panel of Fig.~1 in
\citet{2016arXiv161104954K} shows an integrated form of this function for a
maximum angular separation of $100^\prime$. However, in
\citet{CFHTLenS-2pt-notomo}, the data used in \citet{2016arXiv161104954K},
$\theta$ is between $0.8'$ and $350'$.} of $\xi_\pm$ for $\theta=100'$ and
$\theta=350'$. The lower middle panel in Figure~\ref{fig:filters} shows the
COSEBIs integrands for two angular ranges, $[1',100']$ and $[0.8',350']$, where
we only show the integrands for the lowest COSEBIs mode, $E_1$, as the higher
modes generally probe larger $\ell$-modes.  Finally, the lowest panel shows the
integrands of aperture mass dispersion statistics, for the same two maximum
angular ranges.
